---
title: "Data Stucture Assignment 3"
author: "Jing Pang, Tian Xue, Chuqian Ma, Jiaxiang Leng"
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
---

In the beginning, we load our data into R for further analysis.

```{r}
retention <- read.delim("~/R-workspace/Retention.txt")
```

To be understanding the data pattern, we use summary command code to gather our results. It will show all the detail information, including mean, median, maximum and minimum of every column in the "retention" table. 

```{r}
summary(retention)
```


We could use boxplot to identify our outliers for each column easily. To achieve some basic statistics of each attribute, we use boxplot and histogram to reflect the distribution of them. 

Observing the boxplots, each column of data have some outliers, especially in spend and "rejr" columns. This kind of problem requires more data to justify the data accuracy.


```{r}
par(mfrow=c(1, 3)) 
boxplot(retention$spend, main="spend")
boxplot(retention$apret, main="apret")
boxplot(retention$top10, main="top10")
boxplot(retention$rejr, main="rejr")
boxplot(retention$tstsc, main="tstsc")
boxplot(retention$pacc, main="pacc")
boxplot(retention$strat,main="strat")
```

Then, we will look closer, consider to analysis three specific columns: apret, tstsc, and salar in histogram plot.Then, we will look closer, consider to study three particular columns: apret, tstsc, and salar in the histogram plot.

```{r}
hist(retention$apret, 100)
hist(retention$tstsc, 100)
hist(retention$salar, 100)
```


To estimate the relationship in this three characters, we draw a dot plot graph in between each two of them.

```{r}

pairs(~apret+tstsc+salar,data=retention, main="Cross Relationship Of Characters")

```

For further study, we plot a dot graph of "apret" character constructing separately with "tstsc" and "salar" characters. A linear relationship observes from figures. So we consider modifying a linear regression to predict the future outcomes.


```{r}
plot(y=retention$apret, x=retention$tstsc, main="apret ~ tstsc")
lm1 <- lm(apret ~ tstsc, data=retention)
summary(lm1)
abline(lm1,col="red")

par(mfrow=c(2,2))
plot(lm1)

```
```{r}
plot(y=retention$apret, x=retention$salar, main="apret ~ salar")
lm2 <- lm(apret ~ salar, data=retention)
summary(lm2)
abline(lm2,col="red")


par(mfrow=c(2,2))
plot(lm2)

```


Performing "apret" character on both "tstsc" and "salar" characters displays how the linear relationship occurs on all three characters. The predictable regression function happening to be a plane supports our assumption at the beginning.


```{r}

library(scatterplot3d) 
attach(retention)
s3dplot<- scatterplot3d(salar,tstsc,apret)
lm3 <- lm(apret~salar+tstsc)
summary(lm3)
s3dplot$plane3d(lm3)

par(mfrow=c(2,2))
plot(lm3)
```


We also can test the relationship through another method Anova.

```{r}
fit1 <- lm(apret ~ tstsc + salar, data=retention)
fit2 <- lm(apret ~ tstsc, data=retention)
fit3 <- lm(apret ~ salar, data=retention)
anova(fit1, fit2, fit3)


```


Also, we could use the correlation function to find other characters linear correlation. In this part, we test correction in two different methods Spearman and Pearson. If the value is closer to zero, the relationship of the linear correlation between them is weaker, which means they are more independent on each other.


```{r}

cor(retention, method="spearman")
cor(retention, method="pearson")


```






```{r}

pairs(~spend+apret+top10+rejr+tstsc+pacc+strat+salar,data=retention, 
   main="")

```




In the end, we try to use GeNIe digging more statistical understanding from the dataset. This program helps us locate a Bayesian network between each character. The relationship diagram shows on the following.

