---
title: "Data Analytic Term Project"
author: "Jing Pang, Tian Xue, Chuqian Ma, Jiaxiang Leng"
date: ""
output:
  word_document: default
  pdf_document: default
---

## Introduction 

Kaggle, nowdays, is a very popular place for people who is interesting about data analytic and considering to improve their abilities by some real world problems. This project is to analysis a pretty famous compitition project in Kaggle, which is the Titanic problem. 


data extraction, elimination of null and unuseful data, 

mode trainning

result comparation

Extra: xgboost

data re-analysis, if overfitting, cross-validation, if not, add characters


analysis data, eliminate non-clearly relative data, Linear Regression, prediction mode; Logistic regression and prediction mode on same data. 


## Data Preparation and Statistic Analysis

First of all, the whole datasets are loading into R for further analysis. And then, we used summary code to briefly see the statistics of whole datasets. We will try to analysis that is there any potential problems affecting the later building model.

```{r}
library("ggplot2") 
library("ggthemes") 	
library("scales") 
library("rpart") 
library("randomForest") 
library("party") 

train <- read.csv("~/R-workspace/train.csv")
test <- read.csv("~/R-workspace/test.csv")
gender_submission <- read.csv("~/R-workspace/gender_submission.csv")

summary(train)
summary(test)

```

In this case, we observe some interesting facts about datasets.
There are 4 characters that needs improvement.
Age:
both train and test have NA value 

Cabin:
both train and test have NA value

Embarked 
train have NA value

Fare:
it have NA value in test

There are 2 characters that value is confused.
Ticket number is not consistent for all passages.
Fare range is too wide, contains minimum is 0 and maximum is 512


## Model Version One
For the first version of our model, we simply ignore NA or null value for all characters, and only considering 7 characters in this model built, which are Survived, Pclass, Sex, Age, SibSp, Parch, and Fare.

```{r}
trainMod1 <- train[-c(1,4,9,11)]
trainMod1 <- na.omit(trainMod1)

lgmode <- glm(Survived ~ ., data=trainMod1, family=binomial(link='logit'))

res<- test[c(1)]
res$pred<-predict(object =lgmode,newdata=test, type = "response")
res$pred[is.na(res$pred)]=0
res$predd<-ifelse(res$pred>=0.5,1, 0)

```


## Model Version Two

```{r}
m<- matrix(unlist(strsplit(as.character(train$Name), ", ")),ncol=2,byrow=T)
train$FirstName <- m[,1]
train$LastName <- m[,2]




```



ROC and AUC 

```{r}
library(pROC)
f<-table(gender_submission$Survived, res$predd)

roc_curve <- roc(gender_submission$Survived, res$predd)
names(roc_curve)
x <- 1-roc_curve$specificities
y <- roc_curve$sensitivities

plot(roc_curve, print.auc=TRUE, auc.polygon=TRUE, grid=c(0.1, 0.2),
     grid.col=c("green", "red"), max.auc.polygon=TRUE,
     auc.polygon.col="skyblue", print.thres=TRUE)

```


Version 2

```{r}

# create combination 

test$Survived<-NA
combi<-rbind(train,test)

mode(combi$Name)
combi$Name<-as.character(combi$Name)

combi$Title<-sapply(combi$Name, FUN=function(x){strsplit(x,split='[,.]')[[1]][2]})
head(combi$Title)

combi$Title<-sub(' ','',combi$Title)

data.frame(table(combi$Title))

combi$Title[combi$Title %in% c('Mme', 'Mlle')] <- 'Mlle'
combi$Title[combi$Title %in% c('Capt', 'Don', 'Major', 'Sir')] <- 'Sir'
combi$Title[combi$Title %in% c('Dona', 'Lady', 'the Countess', 'Jonkheer')] <- 'Lady'
combi$Title <- factor(combi$Title)

data.frame(table(combi$Title))

ticket.count <- aggregate(combi$Ticket, by = list(combi$Ticket), function(x)sum(!is.na(x)))
combi$TicketCount <- apply(combi, 1, function(x) 
ticket.count[which(ticket.count[, 1] == x['Ticket']), 2]
)

combi$TicketCount <- factor(sapply(combi$TicketCount, function(x) 
ifelse(x > 1, 'Share', 'Unique')))   ## ifelse(test, yes, no)

combi$FamilySize <- combi$SibSp + combi$Parch + 1

combi$Surname <- sapply(combi$Name, FUN=function(x) {strsplit(x, split='[,.]')[[1]][1]})
combi$FamilyID <- paste(as.character(combi$FamilySize), combi$Surname, sep="")
combi$FamilyID[combi$FamilySize <= 2] <- 'Small'
famIDs <- data.frame(table(combi$FamilyID))

famIDs <- famIDs[famIDs$Freq <= 2,]
combi$FamilyID[combi$FamilyID %in% famIDs$Var1] <- 'Small' 

## FamilySize>2 Freq<=2
combi$FamilyID <- factor(combi$FamilyID)

# list all NA value
sapply(combi,function(x) sum(is.na(x)))
sapply(combi,function(x) sum(x==""))

# age fullfil
age.model <- rpart(Age ~ Pclass + Sex + SibSp + Parch + Fare + Embarked + Title + FamilySize, data=combi[!is.na(combi$Age),], method="anova")
combi$Age[is.na(combi$Age)] <- predict(age.model, combi[is.na(combi$Age),])

## Fare 
combi$Fare[is.na(combi$Fare)] <- median(combi$Fare, na.rm=TRUE)

# Embarked only two value
combi$Embarked[is.na(combi$Embarked)] <- 'C'
combi$Embarked<-factor(combi$Embarked)

# use machine learning
train_df <- combi[1:891,]
test_df <- combi[892:1309,]



```


```{r}

library(party)
library(grid)
library(mvtnorm)
library(modeltools)
library(stats4)
library(strucchange)
library(zoo)
library(sandwich)
set.seed(754)

model <- cforest(as.factor(Survived) ~ Pclass + Sex + Age + Fare + Embarked + 
                 Title + FamilySize + TicketCount + FamilyID, data = train_df,controls=cforest_unbiased(ntree=2000, mtry=3))

res <- predict(model, test_df, OOB=TRUE, type = "response")



write.csv(res,"party_res.csv",row.names=F)

```












## Conclusion 