---
title: "INFSCI 2160 DATA MINING - Homework 1"
author: "Jing Pang"
date: "1/8/2019"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

In this assignment, we will do exploratory analysis on the Black Friday data from Kaggle: https://www.kaggle.com/mehdidag/black-friday in order to complete several questions.

First of all, we have to load the dataset into RStudio for preparation of the further study.

```{r blackfriday, echo=TRUE}
blackfriday <- read.csv("~/R-workspace/BlackFriday.csv", header=TRUE)
```

## Question 1

* How many observations are in the dataset? How many features?

```{r}
nrow(blackfriday)
ncol(blackfriday)
```
There is total number of 537577 observations in the dataset, and there is 12 features in total.

* How many nulls are in the dataset?

```{r, echo=TRUE}
sum(is.na(blackfriday))
```

There is 540285 nulls in the dataset.

* Summarize the dataset.

```{r, echo=TRUE}
summary(blackfriday)
```

* Find the min, max, 1st quartile, 3rd quartile, median, and mean of the ‘Product_Category_1’ column.

```{r, echo=TRUE}
summary(blackfriday$Product_Category_1)
```

* What datatype is the ‘Age’ column?

```{r}
class(blackfriday$Age)
```

## Question 2

* Convert the "Marital_Status" column to a factor 

```{r , echo=TRUE}
blackfriday$Marital_Status <- as.factor(blackfriday$Marital_Status)
class(blackfriday$Marital_Status)
```

## Question 3

* Create a histogram of the ‘Purchase’ column using ggplot2

```{r , echo=TRUE}
library(ggplot2)
ggplot(blackfriday, aes(x = Purchase)) + geom_histogram(binwidth = 50)
```

## Question 4

* Create a table to analyze the ‘City_Category’ column.

```{r, echo=TRUE}
table(blackfriday$City_Category)
```

## Question 5

* Filter the dataset where Gender = M and Marital_Status = 1. How many observations are there?

```{r, echo=TRUE}
library(dplyr)
male_marital_data <- blackfriday %>%
  filter(Gender == "M") %>%
  filter(Marital_Status == 1)
summary(male_marital_data)
```

* Make a table of the age column. Which age group has the most observations?

```{r, echo=TRUE}
AgeTable <- table(blackfriday$Age)
AgeFrame <- as.data.frame(AgeTable)
names(AgeFrame) <- c("Age", "Freq")
library(dplyr)
AgeFrame %>% 
  filter(Freq == max(Freq))
```







